#Load packages
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(purrr)
library(readr)
library(forcats)
library(viridis)
library(Rtsne)
library(tibble)
library(showtext)
library(extrafont)
library(pracma)
library(ggrepel)

windowsFonts(Arial = windowsFont("Arial"))
```

#Load 28d Feces
```{r}
#Imported in ng/g
feces_28  <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Feces_28d.csv")

#remove number of pellets 0
  feces_28 <- feces_28[feces_28$Number.of.Fecal.Pellet.. != 0, ]


doses <- c("Naive", "Vehicle", " PFOA 1.5",  " PFOS 1.5")
print(colnames(feces_28))
pfas_cols <- c(
  "PFOA", "PFOS", "PFBA", "PFPeA", "PFHxA", "PFHpA", "PFNA", "PFDA", "PFUdA", "PFDoA", "PFTrDA", "PFTeDA", "HFPO.DA", "DONA","PFBS", "PFPeS", "PFHxS", "PFHpS", "PFNS", "PFDS",
  "X4_2.FTS", "X6_2_FTS", "X8_2.FTS")
mdls_naive_vehicle_166_05pfoa <- c(
  PFOA = 0.151, PFOS = 0.212, PFBA = 0.138, PFPeA = 0.105, PFHxA = 0.061,
  PFHpA = 0.090, PFNA = 0.112, PFDA = 0.087, PFUdA = 0.109, PFDoA = 0.077,
  PFTrDA = 0.127, PFTeDA = 0.062, HFPO.DA =0.154,DONA= 0.157,PFBS = 0.361, PFPeS = 0.214, PFHxS = 0.192, PFHpS = 0.124, PFNS = 0.225, PFDS = 0.140, `X4_2.FTS` = 0.057, `X6_2_FTS` = 0.098,
  `X8_2.FTS` = 0.093)

mdls_05pos_1pfoa_15pfoa <- c(
  PFOA = 1.44, PFOS = 3.79, PFBA = 0.866, PFPeA = 0.815, PFHxA = 0.770, PFHpA = 0.835, PFNA = 0.774, PFDA = 0.844, PFUdA = 0.633, PFDoA = 0.266, PFTrDA = 0.622, PFTeDA = 0.877, HFPO.DA =2.081,DONA= 0.725,PFBS = 0.602, PFPeS = 2.42, PFHxS = 1.72, PFHpS = 1.40, PFNS = 2.29, PFDS = 1.73, `X4_2.FTS` = 1.27, `X6_2_FTS` = 1.73,
  `X8_2.FTS` = 1.45)

mdls_1pfos <-  c(
  PFOA = 2.89, PFOS = 7.58, PFBA = 1.73, PFPeA = 1.63, PFHxA = 1.54, PFHpA = 1.67, PFNA = 1.55, PFDA = 1.69, PFUdA = 1.27, PFDoA = 0.53, PFTrDA = 1.24, PFTeDA = 1.75, HFPO.DA =4.16,DONA= 1.45,PFBS = 1.20, PFPeS = 4.84, PFHxS = 3.43, PFHpS = 2.79, PFNS = 4.59, PFDS = 3.46, `X4_2.FTS` = 2.54, `X6_2_FTS` = 3.47,
  `X8_2.FTS` = 7.24)

mdls_1.5pfos <-  c(
  PFOA = 7.22, PFOS = 18.95, PFBA = 4.33, PFPeA = 4.08, PFHxA = 3.85, PFHpA = 4.17, PFNA = 3.87, PFDA = 4.22, PFUdA = 3.16, PFDoA = 1.33, PFTrDA = 3.11, PFTeDA = 4.39, HFPO.DA =10.41,DONA= 3.63,PFBS = 3.01, PFPeS = 12.09, PFHxS = 8.58, PFHpS = 6.98, PFNS = 11.47, PFDS = 8.66, `X4_2.FTS` = 6.35, `X6_2_FTS` = 8.67,
  `X8_2.FTS` = 2.90)

mdls_df <- bind_rows(
  tibble(Group = "Naive", !!!mdls_naive_vehicle_166_05pfoa),
  tibble(Group = "Vehicle", !!!mdls_naive_vehicle_166_05pfoa),
  tibble(Group = "PFOA 1.5", !!!mdls_05pos_1pfoa_15pfoa),
  tibble(Group = "PFOS 1.5", !!!mdls_1.5pfos)
) %>%
  pivot_longer(
    -Group,
    names_to = "PFAS",
    values_to = "MDL"
  )

feces_long <- feces_28 %>%
  pivot_longer(
    cols = all_of(pfas_cols),
    names_to = "PFAS",
    values_to = "Value"
  )

feces_long <- feces_long %>%
  left_join(mdls_df, by = c("Group", "PFAS"))

feces_long <- feces_long %>%
  mutate(
    Value = ifelse(Value == "<MDL" & !is.na(MDL),
                   MDL / 2,
                   as.numeric(Value))
  )
feces_28_fixed <- feces_long %>%
  select(-MDL) %>%
  pivot_wider(
    names_from = PFAS,
    values_from = Value
  )


write.csv(feces_28_fixed, file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Feces_PFAS_values_fixed_28d.csv", row.names = FALSE)

```

##Plotting 28d Feces Data
```{r}
feces_28 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Feces_PFAS_values_fixed_28d.csv")

other_pfas <- setdiff(pfas_cols, c("PFOS", "PFOA"))
plasma_colors <- viridis(length(other_pfas), option = "plasma")
names(plasma_colors) <- other_pfas

custom_colors <- c(
  "PFOS" = "red",   
  "PFOA" = "blue",    
  plasma_colors          
)

group_levels <- c(
  "Naive", "Vehicle",
  paste("PFOS", c("1.5")),
  paste("PFOA", c("1.5"))
)

data_combined_long <- feces_28 %>%
  pivot_longer(
    cols = all_of(pfas_cols),
    names_to = "PFAS_name",
    values_to = "PFAS_conc"
  ) %>%
  filter(!is.na(PFAS_conc)) %>%
  mutate(
    PFAS_conc = as.numeric(PFAS_conc),
    group = factor(Group, levels = group_levels),
    PFAS_name = factor(PFAS_name, levels = names(custom_colors)) 
  )
average_pfas <- feces_28 %>%
  group_by(Group) %>%
  summarise(
    Mean_PFOS = format(mean(PFOS, na.rm = TRUE), scientific = FALSE),
    Mean_PFOA = format(mean(PFOA, na.rm = TRUE), scientific = FALSE),
    .groups = "drop"
  )

print(average_pfas)
shape_map <- c(`Male` = 24, `Female` = 21)


data_combined_long <- subset(data_combined_long, !(Group %in% "Naive"))
# Plot
p <- ggplot(data_combined_long, aes(x = group, y = PFAS_conc)) +
  geom_jitter(aes(color = factor(PFAS_name),
                  fill = factor(PFAS_name),
                  shape = Sex),
              size = 2, alpha = 0.5, stroke = 1) +
  scale_color_manual(values = custom_colors, name = "PFAS_name") +
  scale_fill_manual(values = custom_colors, name = "PFAS_name") +
  scale_shape_manual(values = shape_map, name = "Sex") +
 coord_cartesian(ylim = c(0, 15000)) +  
  theme_minimal() +
  theme(
        text = element_text(family = "Arial", size = 12),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      panel.grid.major = element_line(size = 0.5),
      panel.grid.minor = element_blank()
  )
print(p)
ggsave(filename = "../plots/Feces_conc_28d.png", plot = p, width = 5, height = 5, dpi = 300,bg = "white")

replicate_summary <- feces_28 %>%
  group_by(Group) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))

```

#Load 56d Feces
```{r}
#IMPORTED ng/g
feces_56  <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Feces_56d.csv")
#remove number of pellets 0
  feces_56 <- feces_56[feces_56$Number.of.Fecal.Pellet.. != 0, ]

mdls_df <- bind_rows(
  tibble(Group = "Naive", !!!mdls_naive_vehicle_166_05pfoa),
  tibble(Group = "Vehicle", !!!mdls_naive_vehicle_166_05pfoa),
  tibble(Group = "PFOS 0.166", !!!mdls_naive_vehicle_166_05pfoa),
  tibble(Group = "PFOA 0.166", !!!mdls_naive_vehicle_166_05pfoa),
  tibble(Group = "PFOA 0.5", !!!mdls_naive_vehicle_166_05pfoa),
  tibble(Group = "PFOS 0.5", !!!mdls_05pos_1pfoa_15pfoa),
  tibble(Group = "PFOA 1", !!!mdls_05pos_1pfoa_15pfoa),
  tibble(Group = "PFOA 1.5", !!!mdls_05pos_1pfoa_15pfoa),
  tibble(Group = "PFOS 1", !!!mdls_1pfos),
  tibble(Group = "PFOS 1.5", !!!mdls_1.5pfos)
) %>%
  pivot_longer(
    -Group,
    names_to = "PFAS",
    values_to = "MDL"
  )

feces_long <- feces_56 %>%
  pivot_longer(
    cols = all_of(pfas_cols),
    names_to = "PFAS",
    values_to = "Value"
  )

feces_long <- feces_long %>%
  left_join(mdls_df, by = c("Group", "PFAS"))

feces_long <- feces_long %>%
  mutate(
    Value = ifelse(Value == "<MDL" & !is.na(MDL),
                   MDL / 2,
                   as.numeric(Value))
  )
feces_56_fixed <- feces_long %>%
  select(-MDL) %>%
  pivot_wider(
    names_from = PFAS,
    values_from = Value
  )

write.csv(feces_56_fixed, file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Feces_PFAS_values_fixed_56d.csv", row.names = FALSE)
```

##Plotting 56d Feces Data
```{r}
feces_56 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Feces_PFAS_values_fixed_56d.csv" )
pfas_cols <- c(
  "PFOA", "PFOS", "PFBA", "PFPeA", "PFHxA", "PFHpA", "PFNA", "PFDA", "PFUdA", "PFDoA", "PFTrDA", "PFTeDA", "HFPO.DA", "DONA","PFBS", "PFPeS", "PFHxS", "PFHpS", "PFNS", "PFDS",
  "X4_2.FTS", "X6_2_FTS", "X8_2.FTS")
other_pfas <- setdiff(pfas_cols, c("PFOS", "PFOA"))
plasma_colors <- viridis(length(other_pfas), option = "plasma")
names(plasma_colors) <- other_pfas

custom_colors <- c(
  "PFOS" = "red",   
  "PFOA" = "blue",   
  plasma_colors          
)

group_levels <- c("Naive", "Vehicle", "PFOS 0.166", "PFOS 0.5", "PFOS 1", "PFOS 1.5", "PFOA 0.166", "PFOA 0.5", "PFOA 1", "PFOA 1.5")


data_combined_long <- feces_56 %>%
  pivot_longer(
    cols = all_of(pfas_cols),
    names_to = "PFAS_name",
    values_to = "PFAS_conc"
  ) %>%
  filter(!is.na(PFAS_conc)) %>%
  mutate(
    PFAS_conc = as.numeric(PFAS_conc),
    group = factor(Group, levels = group_levels),
    PFAS_name = factor(PFAS_name, levels = names(custom_colors)) 
  )

average_pfas <- feces_56 %>%
  group_by(Group) %>%
  summarise(
    Mean_PFOS = format(mean(PFOS, na.rm = TRUE), scientific = FALSE),
    Mean_PFOA = format(mean(PFOA, na.rm = TRUE), scientific = FALSE),
    .groups = "drop"
  )

print(average_pfas)

shape_map <- c(`Male` = 24, `Female` = 21)
data_combined_long <- subset(data_combined_long, !(Group %in% "Naive"))
# Plot
p <- ggplot(data_combined_long, aes(x = group, y = PFAS_conc)) +
  geom_jitter(aes(color = factor(PFAS_name),
                  fill = factor(PFAS_name),
                  shape = Sex),
              size = 2, alpha = 0.5, stroke = 1) +
  scale_color_manual(values = custom_colors, name = "PFAS_name") +
  scale_fill_manual(values = custom_colors, name = "PFAS_name") +
  scale_shape_manual(values = shape_map, name = "Sex") +
 coord_cartesian(ylim = c(0, 15000)) +  
  theme_minimal() +
  theme(
        text = element_text(family = "Arial", size = 12),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      panel.grid.major = element_line(size = 0.5),
      panel.grid.minor = element_blank()
  )
print(p)
ggsave(filename = "../plots/feces_56_conc.png", plot = p, width = 12.5, height = 5, dpi = 300,bg = "white")

replicate_summary <- feces_56%>%
  group_by(Group) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))
```


#Correlation between liver and feces
```{r}
custom_colors_named <- c(
  "Vehicle" = "#330597",
  "PFOS 0.166"   = "#b12a90",
  "PFOS 0.5"     = "#d35171",
  "PFOS 1"       = "#f68f44",
  "PFOS 1.5"     = "#fec029",
    "PFOA 0.166"   = "#b12a90",
  "PFOA 0.5"     = "#d35171",
  "PFOA 1"       = "#f68f44",
  "PFOA 1.5"     = "#fec029"
)
shape_map <- c(`28` = 24, `56` = 21)
#########56 day
#Imported in mg/kg
liver_56  <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Liver_PFAS_values_fixed_56d.csv")
liver_28  <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Liver_PFAS_values_fixed_28d.csv")

liver_28 <- liver_28 %>% mutate(Time = 28)
liver_56 <- liver_56 %>% mutate(Time = 56)
liver <- rbind(liver_28, liver_56)


liver <- liver %>%
  mutate(Group2 = if_else(Group %in% c("Naive", "Vehicle"),
                         paste("control", Group),
                         Group))
liver <- liver %>%
  separate(Group2, into = c("PFAS", "Dose"), sep = " ", remove = T)

controls <- liver %>%
  filter(Dose %in% c("Naive", "Vehicle")) %>%
  slice(rep(1:n(), each = 2)) %>%
  mutate(PFAS = rep(c("PFOS", "PFOA"), times = n() / 2))

liver <- liver %>%
  filter(!(Dose %in% c("Naive", "Vehicle"))) %>%
  bind_rows(controls)

liver_pfos <- subset(liver) %>%
  select("Mouse.ID", "PFOS", "Group", "Sex", "PFAS", "Time")
liver_pfoa <- subset(liver) %>%
  select("Mouse.ID", "PFOA", "Group", "Sex", "PFAS", "Time")

#imported ng/g
feces_56 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Feces_PFAS_values_fixed_56d.csv" )
feces_28 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Feces_PFAS_values_fixed_28d.csv" )

feces_28 <- feces_28 %>% mutate(Time = 28)
feces_56 <- feces_56 %>% mutate(Time = 56)
feces <- rbind(feces_28, feces_56)

feces <- feces %>% 
  mutate(across(PFOA:X8_2.FTS, ~ .x * 1e-3))  #NG/G TO MG/KG

feces <- feces %>%
  mutate(Group2 = if_else(Group %in% c("Naive", "Vehicle"),
                         paste("control", Group),
                         Group))
feces <- feces %>%
  separate(Group2, into = c("PFAS", "Dose"), sep = " ", remove = T)

controls <- feces %>%
  filter(Dose %in% c("Naive", "Vehicle")) %>%
  slice(rep(1:n(), each = 2)) %>%
  mutate(PFAS = rep(c("PFOS", "PFOA"), times = n() / 2))

feces <- feces %>%
  filter(!(Dose %in% c("Naive", "Vehicle"))) %>%
  bind_rows(controls)

feces_pfos <- subset(feces) %>%
  select("Mouse.ID", "PFOS", "Group", "Sex", "PFAS", "Time")
feces_pfoa <- subset(feces) %>%
  select("Mouse.ID", "PFOA", "Group", "Sex", "PFAS", "Time")

pfos<- inner_join(
  feces_pfos,
  liver_pfos,
  by = c("Mouse.ID","Group", "Sex", "PFAS", "Time"),
  suffix = c("_feces","_liver")
) 
pfos <- subset(pfos, PFAS == "PFOS")

pfoa<- inner_join(
  feces_pfoa,
  liver_pfoa,
  by = c("Mouse.ID","Group", "Sex", "PFAS", "Time"),
  suffix = c( "_feces","_liver")
) 
pfoa <- subset(pfoa, PFAS == "PFOA")

pfos <- subset(pfos, !(Group %in% "Naive"))
pfoa <- subset(pfoa, !(Group %in% "Naive"))
p <- ggplot(pfos, aes(x = PFOS_liver, y = PFOS_feces)) +
  # Single lm line
  geom_smooth(method = "lm", linetype = "dashed", color = "red", inherit.aes = FALSE,
              aes(x = PFOS_liver, y = PFOS_feces)) +
    # Jittered points
  geom_jitter(aes(shape = factor(Time), color = factor(Group), fill = factor(Group)),
              size = 3, alpha = 0.5, stroke = 1) +
  # Labels
  labs(title = "Correlation between Liver and Feces",
       x = "Liver (mg/kg)",
       y = "Feces (mg/kg)") +
  
  # Manual scales
  scale_color_manual(name = "Group", values = custom_colors_named) +
  scale_fill_manual(name = "Group", values = custom_colors_named) +
  scale_shape_manual(name = "Time (days)", values = shape_map) +
  
  # Theme
  theme_minimal() +
  theme(
    text = element_text(family = "Arial", size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major = element_line(size = 0.5),
    panel.grid.minor = element_blank()
  )

p
ggsave(filename = "../plots/fecesxliver_pfos.png", plot = p, width = 7, height = 6, dpi = 300,bg = "white")

p <- ggplot(pfoa, aes(x = PFOA_liver, y = PFOA_feces)) +
  # Single lm line
  geom_smooth(method = "lm", linetype = "dashed", color = "red", inherit.aes = FALSE,
              aes(x = PFOA_liver, y = PFOA_feces)) +
    # Jittered points
  geom_jitter(aes(shape = factor(Time), color = factor(Group), fill = factor(Group)),
              size = 3, alpha = 0.5, stroke = 1) +
  # Labels
  labs(title = "Correlation between Liver and Feces",
       x = "Liver (mg/kg)",
       y = "Feces (mg/kg)") +
  
  # Manual scales
  scale_color_manual(name = "Group", values = custom_colors_named) +
  scale_fill_manual(name = "Group", values = custom_colors_named) +
  scale_shape_manual(name = "Time (days)", values = shape_map) +
  
  # Theme
  theme_minimal() +
  theme(
    text = element_text(family = "Arial", size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major = element_line(size = 0.5),
    panel.grid.minor = element_blank()
  )
p
ggsave(filename = "../plots/fecesxliver_pfoa.png", plot = p, width = 7, height = 6, dpi = 300,bg = "white")
replicate_summary <- pfoa%>%
  group_by(Group, Time) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))


replicate_summary <- pfos%>%
  group_by(Group, Time) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))
```

#Correlation between plasma and feces
```{r}
#########56 day
#Imported in mg/kg
plasma_56 <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Plasma_PFAS_values_fixed_56d.csv")
plasma_28 <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/plasma_PFAS_values_fixed_28d.csv")
plasma_28 <- plasma_28 %>% mutate(Time = 28)
plasma_56 <- plasma_56 %>% mutate(Time = 56)
plasma <- rbind(plasma_28, plasma_56)


plasma <- plasma %>%
  mutate(Group2 = if_else(Group %in% c("Naive", "Vehicle"),
                         paste("control", Group),
                         Group))
plasma <- plasma %>%
  separate(Group2, into = c("PFAS", "Dose"), sep = " ", remove = T)

controls <- plasma %>%
  filter(Dose %in% c("Naive", "Vehicle")) %>%
  slice(rep(1:n(), each = 2)) %>%
  mutate(PFAS = rep(c("PFOS", "PFOA"), times = n() / 2))

plasma <- plasma %>%
  filter(!(Dose %in% c("Naive", "Vehicle"))) %>%
  bind_rows(controls)

plasma_pfos <- subset(plasma) %>%
  select("Mouse.ID", "PFOS", "Group", "Sex", "PFAS", "Time")
plasma_pfoa <- subset(plasma) %>%
  select("Mouse.ID", "PFOA", "Group", "Sex", "PFAS", "Time")

pfos<- inner_join(
  feces_pfos,
  plasma_pfos,
  by = c("Mouse.ID","Group", "Sex", "PFAS", "Time"),
  suffix = c("_feces","_plasma")
) 
pfos <- subset(pfos, PFAS == "PFOS")

pfoa<- inner_join(
  feces_pfoa,
  plasma_pfoa,
  by = c("Mouse.ID","Group", "Sex", "PFAS", "Time"),
  suffix = c( "_feces","_plasma")
) 
pfoa <- subset(pfoa, PFAS == "PFOA")

pfos <- subset(pfos, !(Group %in% "Naive"))
pfoa <- subset(pfoa, !(Group %in% "Naive"))

p <- ggplot(pfos, aes(x = PFOS_plasma, y = PFOS_feces)) +
  # Single lm line
  geom_smooth(method = "lm", linetype = "dashed", color = "red", inherit.aes = FALSE,
              aes(x = PFOS_plasma, y = PFOS_feces)) +
    # Jittered points
  geom_jitter(aes(shape = factor(Time), color = factor(Group), fill = factor(Group)),
              size = 3, alpha = 0.5, stroke = 1) +
  # Labels
  labs(title = "Correlation between plasma and Feces",
       x = "plasma (mg/L)",
       y = "Feces (mg/kg)") +
  
  # Manual scales
  scale_color_manual(name = "Group", values = custom_colors_named) +
  scale_fill_manual(name = "Group", values = custom_colors_named) +
  scale_shape_manual(name = "Time (days)", values = shape_map) +
  
  # Theme
  theme_minimal() +
  theme(
    text = element_text(family = "Arial", size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major = element_line(size = 0.5),
    panel.grid.minor = element_blank()
  )

p
ggsave(filename = "../plots/fecesxplasma_pfos.png", plot = p, width = 7, height = 6, dpi = 300,bg = "white")

p <- ggplot(pfoa, aes(x = PFOA_plasma, y = PFOA_feces)) +
  # Single lm line
  geom_smooth(method = "lm", linetype = "dashed", color = "red", inherit.aes = FALSE,
              aes(x = PFOA_plasma, y = PFOA_feces)) +
    # Jittered points
  geom_jitter(aes(shape = factor(Time), color = factor(Group), fill = factor(Group)),
              size = 3, alpha = 0.5, stroke = 1) +
  # Labels
  labs(title = "Correlation between plasma and Feces",
       x = "plasma (mg/L)",
       y = "Feces (mg/kg)") +
  
  # Manual scales
  scale_color_manual(name = "Group", values = custom_colors_named) +
  scale_fill_manual(name = "Group", values = custom_colors_named) +
  scale_shape_manual(name = "Time (days)", values = shape_map) +
  
  # Theme
  theme_minimal() +
  theme(
    text = element_text(family = "Arial", size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major = element_line(size = 0.5),
    panel.grid.minor = element_blank()
  )
p
ggsave(filename = "../plots/fecesxplasma_pfoa.png", plot = p, width = 7, height = 6, dpi = 300,bg = "white")

replicate_summary <- pfoa%>%
  group_by(Group, Time) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))


replicate_summary <- pfos%>%
  group_by(Group, Time) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))
```

#% of dose feces 
```{r}
df_pfos <- pfos_obs %>%
  filter(Tissue == "Feces", Group != "Vehicle") %>%
  select(Dose, Time, PFAS_conc, Mouse.ID, Sex) %>%
  mutate(
    Chemical = "PFOS",
    ng_kg_feces = PFAS_conc * 1000,
    feces_rate_kg_h_kg = 0.08 / 24,
    ng_h_kg = ng_kg_feces * feces_rate_kg_h_kg,
    Feces_ng_kg = ng_h_kg * 24 * Time,
    Admin_Dose_ng_kg = Dose * Time * 1e6,
    Percent_Dose_Feces = (Feces_ng_kg / Admin_Dose_ng_kg) * 100
  )

## ---- PFOA ----
df_pfoa <- pfoa_obs %>%
  filter(Tissue == "Feces", Group != "Vehicle") %>%
  select(Dose, Time, PFAS_conc, Mouse.ID, Sex) %>%
  mutate(
    Chemical = "PFOA",
    ng_kg_feces = PFAS_conc * 1000,
    feces_rate_kg_h_kg = 0.08 / 24,
    ng_h_kg = ng_kg_feces * feces_rate_kg_h_kg,
    Feces_ng_kg = ng_h_kg * 24 * Time,
    Admin_Dose_ng_kg = Dose * Time * 1e6,
    Percent_Dose_Feces = (Feces_ng_kg / Admin_Dose_ng_kg) * 100
  )

## ---- Combine ----
df_all <- bind_rows(df_pfos, df_pfoa)

shape_map <- c(`Male` = 24, `Female` = 21)
custom_colors_named <- c(
  "PFOS"   = "#f2844b",
  "PFOA"   = "#6a00a8")

df_all$Dose_Time <- paste0(df_all$Dose, " (", df_all$Time, ")")

df_avg <- df_all %>%
  group_by(Chemical, Dose_Time) %>%
  summarise(
    Mean_Percent_Dose_Feces = mean(Percent_Dose_Feces),
    .groups = "drop"
  )

p <- ggplot(df_all,
       aes(
         x = factor(Dose_Time),
         y = Percent_Dose_Feces,
         color = Chemical,
         fill  = Chemical,
         shape = Sex,
         group = interaction(Mouse.ID, Chemical)
       )) +
  geom_jitter(
   size = 2, alpha = 0.5, width = 0.1, stroke = 1
  ) +
  geom_line(
    data = df_avg,
    aes(
      x = factor(Dose_Time),
      y = Mean_Percent_Dose_Feces,
      group = Chemical,
      color = Chemical
    ),
    inherit.aes = FALSE,
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  scale_color_manual(
    name = "Chemical",
    values = custom_colors_named
  ) +
  scale_fill_manual(
    name = "Chemical",
    values = custom_colors_named
  ) +
  scale_shape_manual(
    name = "Sex",
    values = shape_map
  ) +
  
  labs(
    x = NULL,
    y = "% of administered dose in feces"
  ) +
  
  theme_minimal() +
  theme(
    text = element_text(family = "Arial", size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    panel.grid.major = element_line(linewidth = 0.5),
    panel.grid.minor = element_blank()
  )
print(p)
ggsave(filename = "../plots/percent_of_dose.png", plot = p, width = 6, height = 3, dpi = 300,bg = "white")

replicate_summary <- df_all %>%
  group_by(Dose_Time, Chemical) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))

mean <- df_all %>%
  group_by(Chemical) %>%
  summarise(
    n = n(),
    Percent_Dose_Feces = mean(Percent_Dose_Feces, na.rm = TRUE)
  )
```

#Load 28d kidney
```{r}
#Imported in ng/g and i removed the reruns (very similar just used first run)
kidney_28  <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Kidney_28d.csv")

doses <- c("Naive", "Vehicle", " PFOA 1.5",  " PFOS 1.5")

#Load MDLs (so many i had to do it differently this time)
mdls_28  <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/MDLS_Kidney_28d.csv")
chem_cols <- colnames(mdls_28)[4:ncol(mdls_28)]
kidney_28_fixed <-kidney_28

for (chem in chem_cols) {
  
  idx <- grepl("^<", kidney_28_fixed[[chem]])
  
  kidney_28_fixed[[chem]][idx] <- 0.5 * mdls_28[[chem]][idx]
}

write.csv(kidney_28_fixed, file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Kidney_PFAS_values_fixed_28d.csv", row.names = FALSE)

```

##Plotting 28d Kidney Data
```{r}
kidney_28 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Kidney_PFAS_values_fixed_28d.csv")

other_pfas <- setdiff(pfas_cols, c("PFOS", "PFOA"))
plasma_colors <- viridis(length(other_pfas), option = "plasma")
names(plasma_colors) <- other_pfas

custom_colors <- c(
  "PFOS" = "red",   
  "PFOA" = "blue",    
  plasma_colors          
)

group_levels <- c(
  "Naive", "Vehicle",
  paste("PFOS", c("1.5")),
  paste("PFOA", c("1.5"))
)

data_combined_long <- kidney_28 %>%
  pivot_longer(
    cols = all_of(pfas_cols),
    names_to = "PFAS_name",
    values_to = "PFAS_conc"
  ) %>%
  filter(!is.na(PFAS_conc)) %>%
  mutate(
    PFAS_conc = as.numeric(PFAS_conc),
    group = factor(Group, levels = group_levels),
    PFAS_name = factor(PFAS_name, levels = names(custom_colors)) 
  )
average_pfas <- kidney_28 %>%
  group_by(Group) %>%
  summarise(
    Mean_PFOS = format(mean(PFOS, na.rm = TRUE), scientific = FALSE),
    Mean_PFOA = format(mean(PFOA, na.rm = TRUE), scientific = FALSE),
    .groups = "drop"
  )

print(average_pfas)
shape_map <- c(`Male` = 24, `Female` = 21)


data_combined_long <- subset(data_combined_long, !(Group %in% "Naive"))
# Plot
p <- ggplot(data_combined_long, aes(x = group, y = PFAS_conc)) +
  geom_jitter(aes(color = factor(PFAS_name),
                  fill = factor(PFAS_name),
                  shape = Sex),
              size = 2, alpha = 0.5, stroke = 1) +
  scale_color_manual(values = custom_colors, name = "PFAS_name") +
  scale_fill_manual(values = custom_colors, name = "PFAS_name") +
  scale_shape_manual(values = shape_map, name = "Sex") +
 coord_cartesian(ylim = c(0, 150000)) +  
  theme_minimal() +
  theme(
        text = element_text(family = "Arial", size = 12),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      panel.grid.major = element_line(size = 0.5),
      panel.grid.minor = element_blank()
  )
print(p)
ggsave(filename = "../plots/Kidney_28d.png", plot = p, width = 5, height = 5, dpi = 300,bg = "white")

replicate_summary <- kidney_28 %>%
  group_by(Group) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))

```

#Load 56d kidney
```{r}
#IMPORTED ng/g
kidney_56  <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Kidney_56d.csv")

#Load MDLs (so many i had to do it differently this time)
mdls_56  <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/MDLS_Kidney_56d.csv")

chem_cols <- colnames(mdls_56)[4:ncol(mdls_56)]

kidney_56_fixed <-kidney_56

for (chem in chem_cols) {
  
  idx <- grepl("^<", kidney_56_fixed[[chem]])
  
  kidney_56_fixed[[chem]][idx] <- 0.5 * mdls_56[[chem]][idx]
}


write.csv(kidney_56_fixed, file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Kidney_PFAS_values_fixed_56d.csv", row.names = FALSE)
```

##Plotting 56d Kidney Data
```{r}
kidney_56 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Kidney_PFAS_values_fixed_56d.csv" )
pfas_cols <- c(
  "PFOA", "PFOS", "PFBA", "PFPeA", "PFHxA", "PFHpA", "PFNA", "PFDA", "PFUdA", "PFDoA", "PFTrDA", "PFTeDA", "HFPO.DA", "DONA","PFBS", "PFPeS", "PFHxS", "PFHpS", "PFNS", "PFDS",
  "X4_2.FTS", "X6_2_FTS", "X8_2.FTS")
other_pfas <- setdiff(pfas_cols, c("PFOS", "PFOA"))
plasma_colors <- viridis(length(other_pfas), option = "plasma")
names(plasma_colors) <- other_pfas

custom_colors <- c(
  "PFOS" = "red",   
  "PFOA" = "blue",   
  plasma_colors          
)

group_levels <- c("Naive", "Vehicle", "PFOS 0.166", "PFOS 0.5", "PFOS 1", "PFOS 1.5", "PFOA 0.166", "PFOA 0.5", "PFOA 1", "PFOA 1.5")


data_combined_long <- kidney_56 %>%
  pivot_longer(
    cols = all_of(pfas_cols),
    names_to = "PFAS_name",
    values_to = "PFAS_conc"
  ) %>%
  filter(!is.na(PFAS_conc)) %>%
  mutate(
    PFAS_conc = as.numeric(PFAS_conc),
    group = factor(Group, levels = group_levels),
    PFAS_name = factor(PFAS_name, levels = names(custom_colors)) 
  )

average_pfas <- kidney_56 %>%
  group_by(Group) %>%
  summarise(
    Mean_PFOS = format(mean(PFOS, na.rm = TRUE), scientific = FALSE),
    Mean_PFOA = format(mean(PFOA, na.rm = TRUE), scientific = FALSE),
    .groups = "drop"
  )

print(average_pfas)

shape_map <- c(`Male` = 24, `Female` = 21)
data_combined_long <- subset(data_combined_long, !(Group %in% "Naive"))
# Plot
p <- ggplot(data_combined_long, aes(x = group, y = PFAS_conc)) +
  geom_jitter(aes(color = factor(PFAS_name),
                  fill = factor(PFAS_name),
                  shape = Sex),
              size = 2, alpha = 0.5, stroke = 1) +
  scale_color_manual(values = custom_colors, name = "PFAS_name") +
  scale_fill_manual(values = custom_colors, name = "PFAS_name") +
  scale_shape_manual(values = shape_map, name = "Sex") +
 coord_cartesian(ylim = c(0, 150000)) +  
  theme_minimal() +
  theme(
        text = element_text(family = "Arial", size = 12),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      panel.grid.major = element_line(size = 0.5),
      panel.grid.minor = element_blank()
  )
print(p)
ggsave(filename = "../plots/kidney_56_conc.png", plot = p, width = 12.5, height = 5, dpi = 300,bg = "white")

replicate_summary <- kidney_56%>%
  group_by(Group) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))
```

#Correlation between plasma and kidney
```{r}
#########56 day
#Imported in mg/kg
plasma_56 <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Plasma_PFAS_values_fixed_56d.csv")
plasma_28 <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/plasma_PFAS_values_fixed_28d.csv")
plasma_28 <- plasma_28 %>% mutate(Time = 28)
plasma_56 <- plasma_56 %>% mutate(Time = 56)
plasma <- rbind(plasma_28, plasma_56)


plasma <- plasma %>%
  mutate(Group2 = if_else(Group %in% c("Naive", "Vehicle"),
                         paste("control", Group),
                         Group))
plasma <- plasma %>%
  separate(Group2, into = c("PFAS", "Dose"), sep = " ", remove = T)

controls <- plasma %>%
  filter(Dose %in% c("Naive", "Vehicle")) %>%
  slice(rep(1:n(), each = 2)) %>%
  mutate(PFAS = rep(c("PFOS", "PFOA"), times = n() / 2))

plasma <- plasma %>%
  filter(!(Dose %in% c("Naive", "Vehicle"))) %>%
  bind_rows(controls)

plasma_pfos <- subset(plasma) %>%
  select("Mouse.ID", "PFOS", "Group", "Sex", "PFAS", "Time")
plasma_pfoa <- subset(plasma) %>%
  select("Mouse.ID", "PFOA", "Group", "Sex", "PFAS", "Time")

#imported ng/g
kidney_56 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Kidney_PFAS_values_fixed_56d.csv" )
kidney_28 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Kidney_PFAS_values_fixed_28d.csv" )

kidney_28 <- kidney_28 %>% mutate(Time = 28)
kidney_56 <- kidney_56 %>% mutate(Time = 56)
kidney <- rbind(kidney_28, kidney_56)

kidney <- kidney %>% 
  mutate(across(PFOA:X8_2.FTS, ~ .x* 1e-3))  #NG/G TO MG/KG

kidney <- kidney %>%
  mutate(Group2 = if_else(Group %in% c("Naive", "Vehicle"),
                         paste("control", Group),
                         Group))
kidney <- kidney %>%
  separate(Group2, into = c("PFAS", "Dose"), sep = " ", remove = T)

controls <- kidney %>%
  filter(Dose %in% c("Naive", "Vehicle")) %>%
  slice(rep(1:n(), each = 2)) %>%
  mutate(PFAS = rep(c("PFOS", "PFOA"), times = n() / 2))

kidney <- kidney %>%
  filter(!(Dose %in% c("Naive", "Vehicle"))) %>%
  bind_rows(controls)

kidney_pfos <- subset(kidney) %>%
  select("Mouse.ID", "PFOS", "Group", "Sex", "PFAS", "Time")
kidney_pfoa <- subset(kidney) %>%
  select("Mouse.ID", "PFOA", "Group", "Sex", "PFAS", "Time")

pfos<- inner_join(
  kidney_pfos,
  plasma_pfos,
  by = c("Mouse.ID","Group", "Sex", "PFAS", "Time"),
  suffix = c("_kidney","_plasma")
) 
pfos <- subset(pfos, PFAS == "PFOS")

pfoa<- inner_join(
  kidney_pfoa,
  plasma_pfoa,
  by = c("Mouse.ID","Group", "Sex", "PFAS", "Time"),
  suffix = c( "_kidney","_plasma")
) 
pfoa <- subset(pfoa, PFAS == "PFOA")

pfos <- subset(pfos, !(Group %in% "Naive"))
pfoa <- subset(pfoa, !(Group %in% "Naive"))

p <- ggplot(pfos, aes(x = PFOS_plasma, y = PFOS_kidney)) +
  # Single lm line
  geom_smooth(method = "lm", linetype = "dashed", color = "red", inherit.aes = FALSE,
              aes(x = PFOS_plasma, y = PFOS_kidney)) +
    # Jittered points
  geom_jitter(aes(shape = factor(Time), color = factor(Group), fill = factor(Group)),
              size = 3, alpha = 0.5, stroke = 1) +
  # Labels
  labs(title = "Correlation between plasma and Kidney",
       x = "plasma (mg/L)",
       y = "kidney (mg/kg)") +
  
  # Manual scales
  scale_color_manual(name = "Group", values = custom_colors_named) +
  scale_fill_manual(name = "Group", values = custom_colors_named) +
  scale_shape_manual(name = "Time (days)", values = shape_map) +
  
  # Theme
  theme_minimal() +
  theme(
    text = element_text(family = "Arial", size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major = element_line(size = 0.5),
    panel.grid.minor = element_blank()
  )

p
ggsave(filename = "../plots/kidneyxplasma_pfos.png", plot = p, width = 7, height = 6, dpi = 300,bg = "white")

p <- ggplot(pfoa, aes(x = PFOA_plasma, y = PFOA_kidney)) +
  # Single lm line
  geom_smooth(method = "lm", linetype = "dashed", color = "red", inherit.aes = FALSE,
              aes(x = PFOA_plasma, y = PFOA_kidney)) +
    # Jittered points
  geom_jitter(aes(shape = factor(Time), color = factor(Group), fill = factor(Group)),
              size = 3, alpha = 0.5, stroke = 1) +
  # Labels
  labs(title = "Correlation between plasma and Kidney",
       x = "plasma (mg/L)",
       y = "kidney (mg/kg)") +
  
  # Manual scales
  scale_color_manual(name = "Group", values = custom_colors_named) +
  scale_fill_manual(name = "Group", values = custom_colors_named) +
  scale_shape_manual(name = "Time (days)", values = shape_map) +
  
  # Theme
  theme_minimal() +
  theme(
    text = element_text(family = "Arial", size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major = element_line(size = 0.5),
    panel.grid.minor = element_blank()
  )
p
ggsave(filename = "../plots/kidneyxplasma_pfoa.png", plot = p, width = 7, height = 6, dpi = 300,bg = "white")

replicate_summary <- pfoa%>%
  group_by(Group, Time) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))


replicate_summary <- pfos%>%
  group_by(Group, Time) %>%  
  summarise(n = n(), .groups = "drop")
min_n <- min(replicate_summary$n)
max_n <- max(replicate_summary$n)
cat(sprintf("n = %d–%d mice/group\n", min_n, max_n))
```

#Load my data
```{r}
mw_pfos <- 500.13  # g/mol for PFOS
mw_pfoa <- 414.07  # g/mol for PFOA

#Imported in mg/L 
plasma_28 <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/plasma_PFAS_values_fixed_28d.csv")
plasma_56 <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Plasma_PFAS_values_fixed_56d.csv")

#Imported in mg/kg
liver_28  <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Liver_PFAS_values_fixed_28d.csv")
liver_56  <- read.csv("C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Input/Liver_PFAS_values_fixed_56d.csv")

#Imported ng/g
feces_56 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Feces_PFAS_values_fixed_56d.csv" )
feces_28 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Feces_PFAS_values_fixed_28d.csv" )


#Imported ng/g
kidney_56 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Kidney_PFAS_values_fixed_56d.csv" )
kidney_28 <- read.csv(file = "C:/Users/aloan/OneDrive - HC-SC PHAC-ASPC/Documents/PBTK-Modelling-PFAS/Output/Kidney_PFAS_values_fixed_28d.csv" )



process_pfas_file <- function(df, tissue, timepoint) {
  df <- df %>% filter(Group != "Naive")
  df_vehicle <- df %>%
    filter(Group == "Vehicle") %>%
    mutate(
      PFAS_type = "PFOA",
      PFAS_conc = PFOA,
      Tissue = tissue,
      Time = timepoint
    )
  df_vehicle_dup <- df_vehicle %>%
    mutate(
      PFAS_type = "PFOS",
      PFAS_conc = PFOS
    )
  df_vehicles_all <- bind_rows(df_vehicle, df_vehicle_dup)
  df_exposed <- df %>%
    filter(Group != "Vehicle") %>%
    mutate(
      PFAS_type = case_when(
        str_starts(Group, "PFOS") ~ "PFOS",
        str_starts(Group, "PFOA") ~ "PFOA",
        TRUE ~ NA_character_
      ),
      PFAS_conc = case_when(
        PFAS_type == "PFOS" ~ PFOS,
        PFAS_type == "PFOA" ~ PFOA,
        TRUE ~ NA_real_
      ),
      Tissue = tissue,
      Time = timepoint
    )
  df_final <- bind_rows(df_vehicles_all, df_exposed) %>%
    dplyr::select(Mouse.ID, Group, PFAS_type, PFAS_conc, Tissue, Time, Sex)
  return(df_final)
}

plasma_28_clean <- process_pfas_file(plasma_28, tissue = "Plasma", timepoint = 28)
plasma_56_clean <- process_pfas_file(plasma_56, tissue = "Plasma", timepoint = 56)
liver_28_clean  <- process_pfas_file(liver_28,  tissue = "Liver",  timepoint = 28)
liver_56_clean  <- process_pfas_file(liver_56,  tissue = "Liver",  timepoint = 56)
feces_28_clean  <- process_pfas_file(feces_28,  tissue = "Feces",  timepoint = 28)
feces_56_clean  <- process_pfas_file(feces_56,  tissue = "Feces",  timepoint = 56)
kidney_28_clean  <- process_pfas_file(kidney_28,  tissue = "Kidney",  timepoint = 28)
kidney_56_clean  <- process_pfas_file(kidney_56,  tissue = "Kidney",  timepoint = 56)

pfas_obs_all <- bind_rows(plasma_28_clean, plasma_56_clean, liver_28_clean, liver_56_clean,feces_28_clean,feces_56_clean, kidney_56_clean, kidney_28_clean)


pfas_obs_all <- pfas_obs_all %>%
  mutate(
    Dose = case_when(
      Group == "Vehicle" ~ 0,
      TRUE ~ as.numeric(str_extract(Group, "\\d+\\.?\\d*"))
    ),
    mw = case_when(
      PFAS_type == "PFOA" ~ mw_pfoa,
      PFAS_type == "PFOS" ~ mw_pfos,
      TRUE ~ NA_real_
    ),
    PFAS_mg_L_eq = case_when(
      Tissue == "Plasma" ~ PFAS_conc,                 # already mg/L
      Tissue == "Liver" ~ PFAS_conc,              # mg/kg ≈ mg/L
      Tissue == "Feces" ~ PFAS_conc * 1e-3,            # ng/g → mg/kg and same # mg/kg ≈ mg/L
      Tissue == "Kidney" ~ PFAS_conc * 1e-3,            # ng/g → mg/kg and same # mg/kg ≈ mg/L
      TRUE ~ NA_real_
    ),

    # ---- CONVERT TO µM ----
    uM = PFAS_mg_L_eq / mw * 1000,

    # Optional reporting columns
    mg_L = ifelse(Tissue == "Plasma", PFAS_conc, NA_real_),
    ng_L = ifelse(Tissue == "Plasma", PFAS_conc * 1e6, NA_real_),
    mg_kg = case_when(
  Tissue == "Liver"  ~ PFAS_conc,            # already mg/kg
  Tissue == "Kidney" ~ PFAS_conc * 1e-3,     # ng/g → mg/kg
  Tissue == "Feces"  ~ PFAS_conc * 1e-3,     # ng/g → mg/kg
  TRUE ~ NA_real_
)
  ) %>%
  select(-mw)
write.csv(pfas_obs_all,"../Output/observed_data.csv")
```

#HED simple EPA calc
```{r}
#HED = average serum concentration (mg/L) × CL, where CL 81 (L/kg bw/day)

pfas_obs_all <- read.csv("../Output/observed_data.csv")
pfos_obs <- pfas_obs_all %>%
  filter(PFAS_type == "PFOS") %>%
  mutate(mg_L = as.numeric(trimws(mg_L)))
#PFOA
pfoa_obs <- pfas_obs_all %>%
  filter(PFAS_type == "PFOA") %>%
  mutate(mg_L = as.numeric(trimws(mg_L)))

# PFOS HED per individual mouse (range)

HED_pfos_individual <- pfos_obs %>%
  filter(Tissue == "Plasma") %>%
  mutate(hed = mg_L * 4.7355e-06	*24	) %>%      # HED per observation
  group_by(Dose, Time, Mouse.ID)          # keep individual mice

HED_pfos_range <- HED_pfos_individual %>%
  summarise(
    hed_mean  = mean(hed, na.rm = TRUE),
    hed_median = median(hed, na.rm = TRUE),
    hed_min   = min(hed, na.rm = TRUE),
    hed_max   = max(hed, na.rm = TRUE),
    hed_q25   = quantile(hed, 0.25, na.rm = TRUE),
    hed_q75   = quantile(hed, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

# PFOA HED per individual mouse (range)

HED_pfoa_individual <- pfoa_obs %>%
  filter(Tissue == "Plasma") %>%
  mutate(hed = mg_L * 5.5615e-06	*24) %>%
  group_by(Dose, Time, Mouse.ID)

HED_pfoa_range <- HED_pfoa_individual %>%
  summarise(
    hed_mean  = mean(hed, na.rm = TRUE),
    hed_median = median(hed, na.rm = TRUE),
    hed_min   = min(hed, na.rm = TRUE),
    hed_max   = max(hed, na.rm = TRUE),
    hed_q25   = quantile(hed, 0.25, na.rm = TRUE),
    hed_q75   = quantile(hed, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

HED_pfoa_individual$Dose_Time <- paste0(HED_pfoa_individual$Dose, " (", HED_pfoa_individual$Time, ")")
HED_pfos_individual$Dose_Time <- paste0(HED_pfos_individual$Dose, " (", HED_pfos_individual$Time, ")")
```
## CHMS/Olsen et al
```{r}
#load in total data in ng/ml 2018-19, total, 3-79
chms <- data.frame(
  Compound = c("PFOS", "PFOA"),
  GeometricMean = c(2.5, 1.2 ),
  CI_Lower = c(2.3, 1.1),
  CI_Upper = c(2.8, 1.3)
)

chms <- chms %>%
  mutate(
    # ng/mL → mg/L
    C_plasma_mg_L = GeometricMean * 1e-3,

    CL_L_kg_day = case_when(
      # imported as L/h/KG 
      Compound == "PFOS" ~ 4.7355e-06	*24	, 
      Compound == "PFOA" ~ 5.5615e-06	*24
    ),

    Intake_mg_kg_day = C_plasma_mg_L * CL_L_kg_day,

    Intake_low = CI_Lower * 1e-3 * CL_L_kg_day,
    Intake_high = CI_Upper * 1e-3 * CL_L_kg_day
  )

#https://pmc.ncbi.nlm.nih.gov/articles/PMC1964923/
#imported as ng/ml  arithmetric mean 
olsen <- data.frame(
  Compound = c("PFOS", "PFOA"),
  GeometricMean = c(799, 691),
  CI_Lower = c(145, 72),
  CI_Upper = c(3490, 5200)
)
olsen <- olsen %>%
  mutate(
    # ng/mL → mg/L
    C_plasma_mg_L = GeometricMean * 1e-3,

    CL_L_kg_day = case_when(
      # imported as L/h/KG 
      Compound == "PFOS" ~ 4.7355e-06	*24	, 
      Compound == "PFOA" ~ 5.5615e-06	*24
    ),

    Intake_mg_kg_day = C_plasma_mg_L * CL_L_kg_day,

    Intake_low = CI_Lower * 1e-3 * CL_L_kg_day,
    Intake_high = CI_Upper * 1e-3 * CL_L_kg_day
  )
combined <- rbind(olsen, chms)

```


##PFOS Oral equiv
```{r}
pfos_biodata <- combined %>%
  filter(Compound == "PFOS") 
shape_map <- c(`Male` = 24, `Female` = 21)
custom_colors_named <- c(
  "0"   = "#330597",
  "0.166"   = "#b12a90",
  "0.5"     = "#d35171",
  "1"       = "#f68f44",
  "1.5"     = "#fec029")

p <- ggplot() +
  geom_hline(
  data = pfos_biodata,
  aes(yintercept = Intake_mg_kg_day),
  color = c("red", "blue"),
  linetype = "dashed",
  linewidth = 0.8
)+
  # Horizontal lines: colored by Source
geom_rect(
  data = pfos_biodata,
  aes(
    ymin = Intake_low,
    ymax = Intake_high
  ),
  xmin = -Inf,
  xmax = Inf,
  fill = c("red", "blue"),
  alpha = 0.4
) +
  # Points: colored by Dose
  geom_jitter(
    data = HED_pfos_individual, 
    aes(
      x = factor(Dose_Time),
      y = hed,
      shape =Sex,
      color = factor(Dose),
      fill  = factor(Dose),
    ), alpha = 0.6, size = 3, stroke = 1
  ) +

  # Log10 y-axis
scale_y_log10(
  limits = c(1e-8, 1e0),
  breaks = 10^seq(-8, 0, by = 2)
) +
  # Point colors for Dose
  scale_color_manual(
    name   = "Dose (mg/kg/day)",
    values = custom_colors_named
  ) +
  scale_fill_manual(
  name = "Dose",
  values = custom_colors_named
  ) +
    scale_shape_manual(
    name = "Sex",
    values = shape_map
  ) +
  labs(
    title = "Predicted vs Observed PFOS",
    x = "Dose (mg/kg/day)",
    y = "mg/kg/day"
  ) +

  theme_minimal() +
  theme(
    text = element_text(family = "Arial", size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    panel.grid.major = element_line(linewidth = 0.4),
    panel.grid.minor = element_blank()
  )

p

ggsave(filename = "../plots/pfos_hed.png", plot = p, width = 10, height = 4, dpi = 300,bg = "white")

hed_summary_RD <- HED_pfos_individual %>%
  group_by(Dose, Time) %>%
  summarise(
    mean = mean(hed, na.rm = TRUE),
    median = median(hed, na.rm = TRUE),
    min = min(hed, na.rm = TRUE),
    max = max(hed, na.rm = TRUE),
    q25 = quantile(hed, 0.25, na.rm = TRUE),
    q75 = quantile(hed, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

```

##PFOA Oral equiv
```{r}
pfoa_biodata <- combined %>%
  filter(Compound == "PFOA") 
shape_map <- c(`Male` = 24, `Female` = 21)
custom_colors_named <- c(
  "0"   = "#330597",
  "0.166"   = "#b12a90",
  "0.5"     = "#d35171",
  "1"       = "#f68f44",
  "1.5"     = "#fec029")

p <- ggplot() +
  # Points: colored by Dose
  geom_hline(
  data = pfoa_biodata,
  aes(yintercept = Intake_mg_kg_day),
  color = c("red", "blue"),
  linetype = "dashed",
  linewidth = 0.8
)+
  # Horizontal lines: colored by Source
geom_rect(
  data = pfoa_biodata,
  aes(
    ymin = Intake_low,
    ymax = Intake_high
  ),
  xmin = -Inf,
  xmax = Inf,
  fill = c("red", "blue"),
  alpha = 0.4
) +
  geom_jitter(
    data = HED_pfoa_individual, 
    aes(
      x = factor(Dose_Time),
      y = hed,
      shape =Sex,
      color = factor(Dose),
      fill  = factor(Dose),
    ), alpha = 0.6, size = 3, stroke = 1
  ) +

  # Log10 y-axis
scale_y_log10(
  limits = c(1e-8, 1e0),
  breaks = 10^seq(-8, 0, by = 2)
) +
  # Point colors for Dose
  scale_color_manual(
    name   = "Dose (mg/kg/day)",
    values = custom_colors_named
  ) +
  scale_fill_manual(
  name = "Dose",
  values = custom_colors_named
  ) +
    scale_shape_manual(
    name = "Sex",
    values = shape_map
  ) +
  labs(
    title = "Predicted vs Observed PFOA",
    x = "Dose (mg/kg/day)",
    y = "mg/kg/day"
  ) +

  theme_minimal() +
  theme(
    text = element_text(family = "Arial", size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    panel.grid.major = element_line(linewidth = 0.4),
    panel.grid.minor = element_blank()
  )

p

ggsave(filename = "../plots/pfoa_hed.png", plot = p, width = 10, height = 4, dpi = 300,bg = "white")

hed_summary_RD <- HED_pfoa_individual %>%
  group_by(Dose, Time) %>%
  summarise(
    mean = mean(hed, na.rm = TRUE),
    median = median(hed, na.rm = TRUE),
    min = min(hed, na.rm = TRUE),
    max = max(hed, na.rm = TRUE),
    q25 = quantile(hed, 0.25, na.rm = TRUE),
    q75 = quantile(hed, 0.75, na.rm = TRUE),
    .groups = "drop"
  )
```